Improvements of chatbot-applications offering Cognitive Behavioural Therapy regarding Efficacy, Privacy and Safety

### 2021-07-19 paper.tex
further improvements

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
While chatbot-applications in general have become present in areas such as consulting or customer support, they are also implemented
within applications offering Cognitive Behavioural Therapy (CBT). CBT aims to improve a patients mood by changing behavioural patterns and proposing
activities that can help to achieve a higher mindfulness.
\\
Two state-of-the-art chatbots have been reviewed regarding aspects of safety, privacy and efficacy. The review has shown, that they have a positive 
impact on a users mood, take care of sensitive data and can react to critical situations but also have several weaknesses that should be addressed.
\\
As chatbots for mental health belong to Human Computer Interaction (HCI), certain requirements have been reviewed that an application in this field should meet. This review has shown
that users trust high-quality chatbots with sophisticated responses more while having less privacy concerns at the same time. Furthermore, the deletion and 
inappropriate use of data are the two aspects, users worry about the most. 
\\
Based on the review of the existing chatbots and the requirements, three development approaches have been proposed who aim to meet them.
The Knowledge and Data Dashboard, that constantly allows access to saved data and offers options like data deletion, 
the Chatbot Setup Phase trying to personalize the application as much as possible to improve the general user experience and 
a privacy certificate, that ensures the appropriate handling of critical situations.   
\end{abstract}

@@

as non-chatbot based approaches, they can deliver an artificial human interaction that is supposed to increase the

@@

The review and analysis will consider the following points:

@@

  The usage of chatbots for mental health only serves its purpose if the offered therapy is efficiant. Furthermore, recent studies have already examined the efficacy of mental health chatbots.\cite{Fitzpatrick2017,Inkster}

@@

@@

\section{Strenghts and weaknesses of existing chatbots}
\label{sec:two}
This far, there are a couple of chatbots available. This paper will mainly focus on the \emph{Woebot} Chatbot and \emph{Wysa} for several reasons.
On the one hand, they have been developed in a research oriented context. 
The examples are available in english language and free to use.\footnote{https://woebothealth.com/about-us/, last access: 01.05.2021}\textsuperscript{,}\footnote{https://www.wysa.io/faq, last access: 01.05.2021}
\\
Additionally, the developers provide numerous information on their website
like experience reports or technical aspects.
On the other hand, \emph{Wysa} and \emph{Woebot} have been considered in most of recent research.
It revealed some weaknesses of those chatbots, some chances and also risks concerning efficacy, privacy and safety.

@@

a \emph{Facebook} chatbot. It is based on natural language processing and offers free-text inputs as well as predefined answers.

@@

Currently, the usage of \emph{Woebot} is completely free of charge. Nevertheless, \emph{Woebot} Health announces on its website, that they are moving forward

@@

@@

Those proposals of non-technology-based activites are part of evidence-based recommendations for mental health app development\cite{Bakker2016}.
Considering emergency situations, \emph{Woebot} is able to recognize user inputs e.g 'SOS' or 'suicide' as such. It reacts with a list of helplines

@@

free-text input or predefined answers. The biggest difference is the option of having contact to 

@@

The company is located in Bangalore and has been found in collaboration with a research team from the universities of Cambridge and Columbia.

@@

with a less frequent usage. Their participants were human-beings around the globe who have been asked to install the Wysa app. One part of them used the app less often than the other one. In general, both groups showed an improvement. 74\% of the study participants provided an in-app feedback. Most of them preferred the pre-defined answers rather than free-text.

@@

This Section will define requirements a chatbot should meet regarding efficacy, privacy and safety. Those three aspects have already been examined by Kretzschmar et.al \cite{Kretzschmar2019}, but only from the perspection of young adults as a result of a group discussion.

@@

Due to the fact that the usage of a mental health chatbot belongs to HCI, the following requirements will be based on its principles and research conducted in the field.

@@

  \caption{The connection between information, comprehension, control, consciousness and consent, Graphic by Moritz Pflügner, 2021}

@@

the infered knowledge can be seen as information.

@@

The comparison between a physical robot and a chatbot is limited, but the distinguishing betwen data and information can be applied to chatbots as well. It is necessary for chatbot users to know, what inferences it can make based on the conversational text inputs they are providing. The underlying artifical intelligence a chatbot has should be transparently explained 

@@

Ludden et. al \cite{Ludden2015} present personalization as one approach. Considering the PERMA abbrevation, personalization encourage the engagement of a user as the application has a higher personal relevance. If it is used by multiple groups of people, it should be adaptable to them in order to improve the experience.

@@

According to Ischen et.al \cite{Ischen} and Suganuma et.al \cite{Suganuma2018}, services with a higher perception in anthropomorphism lead to a better adherence.

@@

This result corresponds to the results of Ludden et.al \cite{Ludden2015} and hence supports the recommendation to implement the PERMA principles in applications for mental health like chatbots. 

@@

In order to respect the privacy and take user concerns serious, chatbot-applications should only exist as a standlone app without the possibility to use it through a third-party application such as \emph{Facebook}'s messenger in the long term. It is understandable that small companies with limited

@@

This approach could have two major advantages. On the one hand, the chatbot user would always be able to get an insight into data that has been gathered. Additionally, those data could be enriched with the chat snippet or user input that has been the source of it. 

@@

Considering the distinguishing between data and information, questions like 'How did you come to the conclusion that I like rain?' could result in a detailed explaination on how the chatbot infered this particular knowledge.

@@

It would improve the comprehension of the system, giving a human-being control over the data collected by the chatbot and would finally lead to a higher consciousness on what kind of personal data is stored.

@@

Therefore, personalization concerning the user age might be useful. Differences between a personalization for younger people and users who are older could be a different set of questions that are asked in order to examine the current mood.

@@

Addressing people with their name is a quite common personalization approach in current applications and already implemented for Wysa and Woebot. It should be maintained for future developments.

@@

A common approach of chatbot based CBT is the recommendation of certain methods. The chatbot could ask after the accomplishment of a method if the individual liked it or not which would be useful knowledge for further method recommendations. It does not fit in the Chatbot Setup Phase but could

@@

Settings that are infered during the Chatbot-Setup-Phase could be integrated into the Knowledge and Data Dashboard as an extra part. It could be edited there as well if some of the preferences change during the usage.

@@

As Grové \cite{Grove2021} points out, chabots for mental health should notice and react to critical situations in a sophisticated way due to the vulnerable target group. While the reviewed chatbots meet the requirements Kretzschmar et. al \cite{Kretzschmar2019} defined,

@@

With the help of it, clinicians and therapists could easily assess if the application is suitable for their usage in a therapeutical environemnt.\\

@@

that alternative concepts might not have a chance to compete on the market if they do not fulfil a certain aspect. Hence it would be necessary to keep the certificate dynamic. The conditions a chatbot-application has to meet in order to get it should not be static.

\section{Summary}
\label{sec:five}
While the amount of people who use the internet as first source of information has increased\cite{Wyatt2015}, it nowadays offers many applications for CBT such as chatbots. 
\\
Current and popular examples like \emph{Wysa} or \emph{Woebot} are available as a standalone app or through third-party applications. Studies have shown that the usage of them are efficiant, i.e they can improve the mood of their users significantly\cite{Fitzpatrick2017,Inkster}, but in order to tackle problems such as missing trust in a chatbot or a low therapy adherence, further improvements could be implemented like personalization 
or the provision of more transparency. To achieve that, a chatbot setup phase has been proposed in which specific questions could be asked concerning personal preferences and environment of the user. It could be considered as the preliminary talk prior to the therapy a human therapist conducts.
The chatbot setup phase would lead to an implementation of the PERMA aspects, that Ludden et. al \cite{Ludden2015} proposed and which have been presented in \hyperref[sec:three_two]{Section 3.2}. Those aspects are promising to increase the general well-being of a user and hence the efficacy of the chabot application.
\\\\
The developing companies of both reviewed chatbots transparently explain what data is stored for what purpose.
Nevertheless, the review of general requirements for HCI applications has shown, that more information about data collection should be provided and how it is possible for the user to delete it\cite{Saglam2021}.
Furthermore, it has been found that users are not distinguishing between data and information\cite{Lee2011}.
In order to tackle those aspects, a knowledge and data dashboard has been proposed that constantly provides information about the data the chatbot has collected and what it did derive from it.
The main advantages are the constant access to personal and sensitive data on the one hand. On the other hand, the provisioning of it could simplify and prevent administration processes such as deletion requests as they are done automatically via the dashboard.
\\\\
Considering safety aspects, the reviewed chatbots already meet the requirements Kretzschmar et. al \cite{Kretzschmar2019} proposed. Both of them explain that they are just robots and engage the user to do non-digital activities to prevent over-reliance.
It can be considered as a weakness that the only option to figure out if the chatbot notifies and react to critical situations is a trial and error approach. Currently, the developing companies do not publish related information. In order to improve the transparency and to ease the assessment if an application is suitable for therpeutical use, a certificate has been suggested that would confirm the implementation of certain aspects and policies experts consider as crucial.
This certificate could be expanded to privacy and efficacy features as well which could result in a general quality feature for mental health chatbots.

\section{Future Outlook}
In order to get an idea how the field of mental health chatbots will develop in the future, the product pipeline of \emph{WoebotHealth}\footnote{https://woebothealth.com/products-pipeline/, last access: 31.05.2021} gives an insight in their developments. They are moving forward to implement several chatbot solutions for specific target groups. Besides the presented general one for adult mental health, they also offer a variant of \emph{Woebot} for adolescents, pregnant persons or 
individuals with a problematic substance use.
\\
The approach to split up the general chatbot-application into several specific ones aiming different target groups could be promising to increase the efficacy even further due to therapy methods that are tailored to the patient's situation. 
\\\\
Furthermore \emph{WoebotHealth} is currently developing prescription digital therapeutics who aim to treat symptoms of certain mental illnesses. All of them are currently in the admission process, while the digital therapeutical \emph{WB001} has been designated by the \emph{FDA} (U.S Food and Drug Administration) as \emph{Breakthrough Device}.\textsuperscript{18}
\\
The development of prescription digital therapeutics could be one aspect regarding a more sustainable business model. Not only \emph{WoebotHealth} tries to reach health insurance providers, also \emph{Touchkin} announce on their website\footnote{https://www.wysa.io/, last access: 31.05.2021} that they are preparing to cooperate with them what could increase the number of users.
\\
Additionally they are also offering their services to employers who might have a genuine interest in the mental health of their employees. Using mental health application in workplace environments could play a bigger role in the near future as it is promising to fill the gap between meditation and employee assistance programs.\footnote{https://www.wysa.io/for-employers, last access: 31.05.2021}
\\\\
With the development of the internet of things, it is also conceivable that chatbots for mental health become ubiquitous. Being able to start a conversation constantly, for example with a smart watch, might have a positive impact on the therapy adherence. According to Donkin et. al \cite{Donkin2011}, this would also have a positive impact on the the efficacy of the application in general.
\\
Nevertheless, more research is necessary to examine the connection between ubiquitous computing and the use of mental health chatbots.
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{unsrt}
\bibliography{library}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
While chatbot-applications in general have become present in areas such as consulting or customer support, they are also implemented
within applications offering Cognitive Behavioural Therapy (CBT). CBT aims to improve a patient's mood by changing behavioural patterns and proposing
activities that can help to achieve a higher mindfulness.
\\
Two state-of-the-art chatbots have been reviewed regarding aspects of safety, privacy and efficacy. The review has shown, that they have a positive 
impact on a user's mood, take care of sensitive data and can react to critical situations but also have several weaknesses that should be addressed.
\\
As chatbots for mental health belong to Human Computer Interaction (HCI), certain requirements have been reviewed that an application in this field should meet. This review has shown
that users trust high-quality chatbots with sophisticated responses more while having less privacy concerns at the same time. Furthermore, the deletion and 
inappropriate use of data are the two aspects, users worry about the most. 
\\
Based on the review of the existing chatbots and the requirements, three development approaches have been proposed who aim to meet them:
The Knowledge and Data Dashboard, that constantly allows access to saved data and offers options like data deletion, 
the Chatbot Setup Phase trying to personalize the application as much as possible to improve the general user experience and 
a privacy certificate, that ensures the appropriate handling of critical situations.   
\end{abstract}

@@

as human-based approaches, they can deliver an artificial human interaction that is supposed to increase the

@@

The review and analysis will consider the following points. 

@@

  The usage of chatbots for mental health only serves its purpose if the offered therapy is efficient. Furthermore, recent studies have already examined the efficacy of mental health chatbots.\cite{Fitzpatrick2017,Inkster}

@@

On the one hand, at least one aspect out of them has been considered in most of the referenced literature. On the other hand, Kretzschmar et. al \cite{Kretzschmar2019} found out, that users of mental health chatbots consider 
them as most important.

@@

\section{Strenghts and weaknesses of existing chatbots}
\label{sec:two}
This far, there are a couple of chatbots available. This paper will mainly focus on the \emph{Woebot} Chatbot and \emph{Wysa} for several reasons.
They are the only apps that offer a real chatbot. Others like AbleTo\footnote{https://www.ableto.com/} or talkspace\footnote{https://try.talkspace.com/} focus more on online CBT with a real therapist.
On the one hand, \emph{Woebot} and \emph{Wysa} have been developed in a research oriented context. 
The examples are available in english language and free to use.\footnote{https://woebothealth.com/about-us/, last access: 01.05.2021}\textsuperscript{,}\footnote{https://www.wysa.io/faq, last access: 01.05.2021}
Additionally, the developers provide numerous information on their website
like experience reports or technical aspects.
On the other hand, \emph{Wysa} and \emph{Woebot} have been considered in most of recent research.
It revealed some weaknesses of those chatbots, some chances and also risks concerning efficacy, privacy and safety.

@@

a \emph{Facebook} chatbot. It is based on natural language processing and offers free text inputs as well as predefined answers.

@@

Currently, the usage of \emph{Woebot} is completely free of charge. Nevertheless, \emph{WoebotHealth} announces on its website, that they are moving forward

@@

A review of \emph{Woebot} by Megan Molteni published in the WIRED magazine \cite{Molteni2017} critizises the possibilty of accessing the chatbot through the messenger of \emph{Facebook}.
She points out that medical data privacy and security law can not be applied to \emph{Woebot} conversations as it is not considered as a licensed medical provider when accessed through it.

@@

Those proposals of non-technology-based activities are part of evidence-based recommendations for mental health app development\cite{Bakker2016}.
Considering emergency situations, \emph{Woebot} is able to recognize user inputs e.g. 'SOS' or 'suicide' as such. It reacts with a list of helplines

@@

free text input or predefined answers. The biggest difference is the option of having contact to 

@@

The company is located in Bangalore and has been founded in collaboration with a research team from the universities of Cambridge and Columbia.

@@

with a less frequent usage. Their participants were human beings around the globe who have been asked to install the Wysa app. One part of them used the app less often than the other one. In general, both groups showed an improvement. 74\% of the study participants provided an in-app feedback. Most of them preferred the pre-defined answers rather than free text.

@@

This section will define requirements a chatbot should meet regarding efficacy, privacy and safety. Those three aspects have already been examined by Kretzschmar et.al \cite{Kretzschmar2019}, but only from the perspection of young adults as a result of a group discussion.

@@

Due to the fact that the usage of a mental health chatbot belongs to the field of HCI, the following requirements will be based on its principles and research conducted in the field.

@@

  \caption{The connection between information, comprehension, control, consciousness and consent, graphic by Moritz Pflügner, 2021}

@@

the inferred knowledge can be seen as information.

@@

The comparability between a physical robot and a chatbot is limited, but the distinction between data and information can be applied to chatbots as well. It is necessary for chatbot users to know, what inferences the bot can make based on the conversational text inputs they are providing. The underlying artifical intelligence a chatbot has should be transparently explained 

@@

Ludden et. al \cite{Ludden2015} present personalization as one approach. Considering the PERMA abbrevation, personalization encourages the engagement of a user as the application has a higher personal relevance. If it is used by multiple groups of people, it should be adaptable to them in order to improve the experience.

@@

According to Ischen et.al \cite{Ischen} and Suganuma et.al \cite{Suganuma2018}, services with a higher perception of anthropomorphism lead to a better adherence.

@@

This result corresponds to the results of Ludden et al. \cite{Ludden2015} and hence supports the recommendation to implement the PERMA principles in applications for mental health like chatbots. 

@@

\begin{figure*}[h]
  \includegraphics[width=\textwidth]{kddb}
  \caption{design mockup of a knowledge and data dashboard, graphic by Moritz Pflügner, 2021}
  \label{fig:kddb}
\end{figure*}
In order to respect the privacy and take user concerns seriously, chatbot-applications should only exist as a standalone app without the possibility to use it through a third-party application such as \emph{Facebook}'s messenger in the long term. It is understandable that small companies with limited

@@

This approach could have two major advantages. On the one hand, the chatbot user would always be able to get an insight into data that has been gathered. Additionally, the data could be enriched with the chat snippet or user input that has been the source of it. 

@@

Considering the distinction between data and information, questions like 'How did you come to the conclusion that I like rain?' could result in a detailed explaination on how the chatbot inferred this particular knowledge.

@@

It would improve the comprehension of the system, giving a human being control over the data collected by the chatbot and would finally lead to a higher consciousness on what kind of personal data is stored.

@@

Therefore, personalization concerning the user's age might be useful. Differences between a personalization for younger people and users who are older could be a different set of questions that are asked in order to examine the current mood.

@@

Addressing people with their name is quite a common personalization approach in current applications and already implemented for Wysa and Woebot. It should be maintained for future developments.

@@

A common approach of chatbot based CBT is the recommendation of certain CBT methods. The chatbot could ask after the accomplishment of a method if the individual liked it or not which would be useful knowledge for further method recommendations. It does not fit in the Chatbot Setup Phase but could

@@

  \includegraphics[width=\linewidth]{chatbot_setup_phase}
  \caption{design mockup of a chatbot setup phase, graphic by Moritz Pflügner, 2021}
  \label{fig:setup_phase}
Settings that are inferred during the Chatbot-Setup-Phase could be integrated into the Knowledge and Data Dashboard as an extra part. It could be edited there as well if some of the preferences change during the usage.

@@

As Grové \cite{Grove2021} points out, chatbots for mental health should notice and react to critical situations in a sophisticated way due to the vulnerable target group. While the reviewed chatbots meet the requirements Kretzschmar et. al \cite{Kretzschmar2019} defined,

@@

With the help of it, clinicians and therapists could easily assess if the application is suitable for their usage in a therapeutical environment.\\

@@

that alternative concepts might not have a chance to compete on the market if they do not fulfil a certain aspect. Hence it would be necessary to keep the certificate dynamic. The conditions a chatbot-application has to meet in order to get it should not be static and regularly re-examined by experts with different backgrounds.
Another approach would be a product-specific assessment. Companies who want to get the certificate would have to explain and justify their concepts and implementation of the mentioned aspects. Based on this presentation, the certificate could be awarded.

\section{Summary}
\label{sec:five}
While the amount of people who use the internet as a first source of information has increased\cite{Wyatt2015}, it nowadays offers many applications for CBT such as chatbots. 
\\
Current and popular examples like \emph{Wysa} or \emph{Woebot} are available as a standalone app or through third-party applications. Studies have shown that the usage of them are efficient, i.e. they can improve the mood of their users significantly\cite{Fitzpatrick2017,Inkster}, but in order to tackle problems such as missing trust in a chatbot or a low therapy adherence, further improvements could be implemented like personalization 
or the provision of more transparency. To achieve that, a chatbot setup phase has been proposed in which specific questions could be asked concerning personal preferences and environment of the user. It could be considered as the preliminary talk prior to the therapy a human therapist conducts.
The chatbot setup phase would lead to an implementation of the PERMA aspects, that Ludden et. al \cite{Ludden2015} proposed and which have been presented in \hyperref[sec:three_two]{Section 3.2}. Those aspects are promising to increase the general well-being of a user and hence the efficacy of the chatbot application.
\\\\
The developing companies of both reviewed chatbots transparently explain what data is stored for what purpose.
Nevertheless, the review of general requirements for HCI applications has shown, that more information about data collection should be provided and how it is possible for the user to delete it\cite{Saglam2021}.
Furthermore, it has been found that users are not distinction between data and information\cite{Lee2011}.
In order to tackle those aspects, a knowledge and data dashboard has been proposed that constantly provides information about the data the chatbot has collected and what it did derive from it.
The main advantages are the constant access to personal and sensitive data on the one hand. On the other hand, the provisioning of it could simplify and prevent administration processes such as deletion requests as they are done automatically via the dashboard.
\\\\
Considering safety aspects, the reviewed chatbots already meet the requirements Kretzschmar et. al \cite{Kretzschmar2019} proposed. Both of them explain that they are just robots and engage the user to do non-digital activities to prevent over-reliance.
It can be considered a weakness that the only option to figure out if the chatbot notices and reacts to critical situations is a trial and error approach. Currently, the developing companies do not publish related information. In order to improve the transparency and to ease the assessment if an application is suitable for therapeutical use, a certificate has been suggested that would confirm the implementation of certain aspects and policies experts consider as crucial.
This certificate could be expanded to privacy and efficacy features as well which could result in a general quality feature for mental health chatbots.

\section{Future Outlook}
In order to get an idea how the field of mental health chatbots will develop in the future, the product pipeline of \emph{WoebotHealth}\footnote{https://woebothealth.com/products-pipeline/, last access: 31.05.2021} gives an insight in their developments. They are moving forward to implement several chatbot solutions for specific target groups. Besides the presented general one for adult mental health, they also offer a variant of \emph{Woebot} for adolescents, pregnant persons or 
individuals with a problematic substance use.
\\
The approach to split up the general chatbot-application into several specific ones aiming different target groups could be promising to increase the efficacy even further due to therapy methods that are tailored to the patient's situation. 
\\\\
Furthermore \emph{WoebotHealth} is currently developing prescription digital therapeutics who aim to treat symptoms of certain mental illnesses. All of them are currently in the admission process, while the digital therapeutical \emph{WB001} has been designated by the \emph{FDA} (U.S Food and Drug Administration) as \emph{Breakthrough Device}.\textsuperscript{18}
\\
The development of prescription digital therapeutics could be one aspect regarding a more sustainable business model. Not only \emph{WoebotHealth} tries to reach health insurance providers, also \emph{Touchkin} announce on their website\footnote{https://www.wysa.io/, last access: 31.05.2021} that they are preparing to cooperate with them which could increase the number of users.
\\
Additionally they are also offering their services to employers who might have a genuine interest in the mental health of their employees. Using mental health application in workplace environments could play a bigger role in the near future as it is promising to fill the gap between meditation and employee assistance programs.\footnote{https://www.wysa.io/for-employers, last access: 31.05.2021}
\\\\
With the development of the internet of things, it is also conceivable that chatbots for mental health become ubiquitous. Being able to start a conversation at any moment, for example with a smart watch, might have a positive impact on the therapy adherence. According to Donkin et. al \cite{Donkin2011}, this would also have a positive impact on the the efficacy of the application in general.
\\
Nevertheless, more research is necessary to examine the connection between ubiquitous computing and the use of mental health chatbots.
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{unsrt}
\bibliography{library}
